Bayesian Rate Estimation

### 2012-08-09 many.tex
Fixed URL typo.

\author{Ilya Mandel}
\email{imandel@star.sr.bham.ac.uk}
\homepage{http://www.star.sr.bham.ac.uk/~imandel}
\affiliation{School of Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT Birmingham\\United Kingdom}
%\affiliation{University of Birmingham}

@@

We assume that both the foreground and background events are samples
from an inhomogeneous \ilya{[pardon my ignorance, what does "inhomogeneous" refer to in this context (i.e., what would a homogeneous Poisson process be?)]} Poisson process with \ilya{respective} differential rates
\begin{equation}
  \frac{dN_f}{dx} = f(x, \theta)
\end{equation}
and 
\begin{equation}
  \frac{dN_b}{dx} = b(x, \theta),
\end{equation}
where the $\theta$ argument represents additional ``shape'' parameters
that may affect the distribution, and for which we will eventually
fit.  The cumulative rates of the two processes are therefore
\begin{equation}
  F(x,\theta) \equiv \int_{-\infty}^x ds\, f(s, \theta)
\end{equation}
and
\begin{equation}
  B(x,\theta) \equiv \int_{-\infty}^x ds\, b(s,\theta).
\end{equation}
The assumption that the foreground and background events form an
inhomogeneous Poisson process implies
\begin{enumerate}
\item \label{prop:Poisson} The number of events in any range of
  ranking statistics, $x \in [x_1, x_2]$ is Poisson distributed with
  rate $F(x_2, \theta) - F(x_1,\theta)$ or $B(x_2,\theta) -
  B(x_1,\theta)$.
\item The numbers of events in non-overlapping ranges of ranking
  statistics are independent. 
\item The probability of exactly one foreground event between $x$ and
  $x+h$ is given by
  \begin{equation}
    P(\ilya{n} = 1 \in [x, x+h]) = f(x,\theta) h + \order{h^2}.
  \end{equation}
  and similarly for background events.
\item The probability of two or more events in a small range of
  ranking statistic is negligible
  \begin{equation}
    P(\ilya{n} = 2 \in [x, x+h]) = \order{h^2}.
  \end{equation}
\end{enumerate}
The foreground and background rates can in general depend on several
parameters; the goal of our analysis is to determine the posterior
probability distributions for these parameters that are implied by the
data.  At the least, we will want to know the overall amplitude of the
foreground and background rates.  Let
\begin{equation}
  f(x,\theta) = R_f \hat{f}(x,\theta'),
\end{equation}
and 
\begin{equation}
  b(x, \theta) = R_b \hat{b}(x, \theta'),
\end{equation}
where $\hat{F}(\infty, \theta') = \hat{B}(\infty, \theta') = 1$, and
$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f\ilya{\equiv F(\infty,\theta)}$ and $R_b\ilya{\equiv B(\infty,\theta)}$
are the total number of foreground and background events expected and
$\hat{f}(x, \theta')$ and $\hat{b}(x, \theta')$ are the likelihood of
obtaining an event with ranking statistic $x$ under the foreground and
background distributions.  In what follows, we will drop the prime,
using $\theta$ to denote all parameters of the rate distributions
except $R_f$ and $R_b$.

>>>>

\author{Ilya Mandel}
\email{imandel@star.sr.bham.ac.uk}
\homepage{http://www.sr.bham.ac.uk/~imandel}
\affiliation{School of Physics and Astronomy\\University of Birmingham\\Edgbaston B15 2TT Birmingham\\United Kingdom}
%\affiliation{University of Birmingham}

@@

We assume that both the foreground and background events are samples
from an inhomogeneous Poisson process \ilya{[pardon my ignorance, but I thought an "inhomogeneous" Poisson process meant that the rate changed as a function of time -- and I don't think that's what we are assuming...]} with \ilya{respective} differential rates
\begin{equation}
  \frac{dN_f}{dx} = f(x, \theta)
\end{equation}
and 
\begin{equation}
  \frac{dN_b}{dx} = b(x, \theta),
\end{equation}
where the $\theta$ argument represents additional ``shape'' parameters
that may affect the distribution, and for which we will eventually
fit.  The cumulative rates of the two processes are therefore
\begin{equation}
  F(x,\theta) \equiv \int_{-\infty}^x ds\, f(s, \theta)
\end{equation}
and
\begin{equation}
  B(x,\theta) \equiv \int_{-\infty}^x ds\, b(s,\theta).
\end{equation}
The assumption that the foreground and background events form an
inhomogeneous Poisson process implies
\begin{enumerate}
\item \label{prop:Poisson} The number of events in any range of
  ranking statistics, $x \in [x_1, x_2]$ is Poisson distributed with
  rate $F(x_2, \theta) - F(x_1,\theta)$ or $B(x_2,\theta) -
  B(x_1,\theta)$.
\item The numbers of events in non-overlapping ranges of ranking
  statistics are independent. 
\item The probability of exactly one foreground event between $x$ and
  $x+h$ is given by
  \begin{equation}
    P(\ilya{n} = 1 \in [x, x+h]) = f(x,\theta) h + \order{h^2}.
  \end{equation}
  and similarly for background events.
\item The probability of two or more events in a small range of
  ranking statistic is negligible
  \begin{equation}
    P(\ilya{n} = 2 \in [x, x+h]) = \order{h^2}.
  \end{equation}
\end{enumerate}
The foreground and background rates can in general depend on several
parameters; the goal of our analysis is to determine the posterior
probability distributions for these parameters that are implied by the
data.  At the least, we will want to know the overall amplitude of the
foreground and background rates.  Let
\begin{equation}
  f(x,\theta) = R_f \hat{f}(x,\theta'),
\end{equation}
and 
\begin{equation}
  b(x, \theta) = R_b \hat{b}(x, \theta'),
\end{equation}
where $\hat{F}(\infty, \theta') = \hat{B}(\infty, \theta') = 1$, and
$\theta' = \theta \setminus \{R_{f}, R_{b} \}$.  Then $R_f\ilya{\equiv F(\infty,\theta)}$ and $R_b\ilya{\equiv B(\infty,\theta)}$
are the total number of foreground and background events expected and
$\hat{f}(x, \theta')$ and $\hat{b}(x, \theta')$ are the likelihood of
obtaining an event with ranking statistic $x$ under the foreground and
background distributions.  In what follows, we will drop the prime,
using $\theta$ to denote all parameters of the rate distributions
except $R_f$ and $R_b$.
